---
title: "Ceteris Paribus Plots for classification"
author: "Przemyslaw Biecek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Title of your vignette}
  %\usepackage[UTF-8]{inputenc}
---

# Introduction

Here we present an example for classification. 
The `status` variable in the `HR` dataset is a factor that takes three values `fired`/`ok`/`promoted`. We will create a model for this variable and then we present explainers for this model.

```{r, warning=FALSE, message=FALSE}
library("DALEX")
head(HR)
```

Here we create a random forest model for this dataset.

```{r, warning=FALSE, message=FALSE}
library("randomForest")
library("ceterisParibus")
set.seed(59)

model <- randomForest(status ~ gender + age + hours + evaluation + salary, data = HR)
```

By default the `predict.randomForest()` function returns classes not scores. This is why we use user-specific `predict()` function. Here we use two explainers, one will explainer the `fired` class while the second will take care about the `promoted` class.

```{r, warning=FALSE, message=FALSE}
pred1 <- function(m, x)   predict(m, x, type = "prob")[,1]
pred3 <- function(m, x)   predict(m, x, type = "prob")[,3]

explainer_rf_fired <- explain(model, data = HR[,1:5], 
                              y = HR$status == "fired", 
                              predict_function = pred1, label = "fired")
explainer_rf_promoted <- explain(model, data = HR[,1:5], 
                               y = HR$status == "promoted", 
                               predict_function = pred3, label = "promoted")
```

# Individial curves

The `explainer_rf_fired` explainer is focused on class `fired`. Let's see Ceteris Paribus profiles for first 10 individuals.

They are colored with the `gender` variable. It's useful since in the model there is an interaction between age and gender. Can you spot it?

```{r, warning=FALSE, message=FALSE}
cp_rf_fired <- ceteris_paribus(explainer_rf_fired, HR[1:10,])

plot(cp_rf_fired, color = "gender")
```

# Partial Dependency Plots

We can plot explainers for two classes (fired and promoted) in the same plot. Here is an example how to do this.

The `hours` variable is an interesting one. People with low average working hours are on average more likely to be fired not promoted. Also the `salary` seems to be related with probability of being `fired` but not `promoted`.


```{r, warning=FALSE, message=FALSE}
cp_rf1 <- ceteris_paribus(explainer_rf_fired, HR[1:100,])
cp_rf3 <- ceteris_paribus(explainer_rf_promoted, HR[1:100,])

plot(cp_rf1, cp_rf3, 
     aggregate_profiles = mean, 
     alpha = 1, show_observations = FALSE, color="_label_")
```

# Partial Dependency Plots for groups of obervations

Here we will compare the explainer (probability of being fired) across two genders.
We know that there is an interaction between `gender` and `age` variables. It will be easy to read this interaction from the plot.

In the plot below we will see that both genders behave in all panels except age.
For the `age` variable it looks like younger woman are more likely to be fired while for males the older age is a risk factor. 

Is it because employers are afraid of maternity leaves? Maybe, but please note, that this dataset is an artificial/simulated one.

```{r, warning=FALSE, message=FALSE}
cp_rfF <- ceteris_paribus(explainer_rf_fired, 
                          HR[which(HR$gender == "female")[1:100],])
cp_rfF$`_label_` = "Fired Female"
cp_rfM <- ceteris_paribus(explainer_rf_fired, 
                          HR[which(HR$gender == "male")[1:100],])
cp_rfM$`_label_` = "Fired Male"

plot(cp_rfM, cp_rfF, 
     aggregate_profiles = mean, 
     alpha = 1, show_observations = FALSE, color="_label_")
```


# Session info

```{r, warning=FALSE, message=FALSE}
devtools::session_info()
```

